---
title: "DS607_FOMC_Sentiment_Analysis"
author: "Jagdish Chhabria"
date: "April 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load required packages

```{r download statements}
library(rlang)
library(pdftools)
library(tidyverse)
library(stringr)
library(tidytext)
library(tidyr)
library(xlsx)
library(RCurl)
library(XML)
library(kableExtra)
library(tm)
library(ngram)
library(SentimentAnalysis)
library(tidytext)
#library(ggridges)
library(wordcloud)
library(ggplot2)
```

### Collate links to FOMC statements from January 2007 till date (May 2019)

This is one of the primary sources of data. The URL for each FOMC statement from  January 2007 onwards is collated from the following website (https://www.federalreserve.gov/monetarypolicy). While historical statements are available from 1994 onwards, the choice of starting point i.e. January 2007 was made to strike a balance between quantity and quality (adequate representation through different periods of the economic cycle such as recession (2009) and growth (2018).

There is a total of 102 statements published during this period.

```{r}

links<-c("https://www.federalreserve.gov/newsevents/pressreleases/monetary20190130a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20190320a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20190501a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20180131a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20180321a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20180502a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20180613a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20180801a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20180926a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20181108a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20181219a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20170201a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20170315a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20170503a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20170614a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20170726a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20170920a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20171101a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20171213a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20160127a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20160316a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20160427a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20160615a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20160727a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20160921a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20161102a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20161214a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20150128a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20150318a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20150429a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20150617a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20150729a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20150917a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20151028a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20151216a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20140129a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20140319a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20140430a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20140618a.htm", 
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20140730a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20140917a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20141029a.htm",  
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20141217a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20130130a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20130320a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20130501a.htm", 
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20130619a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20130731a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20130918a.htm",  
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20131030a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20131218a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20120125a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20120313a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20120425a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20120620a.htm", 
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20120801a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20120913a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20121024a.htm",  
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20121212a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20110126a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20110315a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20110427a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20110622a.htm", 
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20110809a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20110921a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20111102a.htm",  
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20111213a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20100127a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20100316a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20100428a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20100623a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20100810a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20100921a.htm",  
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20101103a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20101214a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20090128a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20090318a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20090429a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20090624a.htm", 
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20090812a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20090923a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20091104a.htm",  
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20091216a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20080122b.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20080130a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20080318a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20080430a.htm", 
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20080625a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20080805a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20080916a.htm",  
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20081008a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20081029a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20081216b.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20070131a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20070321a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20070509a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20070618a.htm", 
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20070807a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20070817b.htm",  
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20070918a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20071031a.htm",
         "https://www.federalreserve.gov/newsevents/pressreleases/monetary20071211a.htm"
         
)

length(links)


```

## Data Preparation

```{r}

# Extract year of publication from the statement's release date, and create a data frame with date, year and URL. 
# Convert factors to characters

statement.dates<-NULL
year<-NULL
for(i in seq(from=1, to=length(links))) {
  statement.dates[i]<-(str_extract(links[i],"[[:digit:]]+"))
  year[i]<-substr(statement.dates[i],1,4)
}

reports<-data.frame(year,statement.dates, links)
reports %<>% mutate_if(is.factor, as.character)%>% arrange(statement.dates)

```


```{r}
# Loop through the statement links and scrape the content
# Discard irrelevant portions of the extracted content i.e. prelimary paragraphs and last paragraph


statement.content<-NULL
statement.length<-NULL
for(i in seq(from=1, to=length(reports$links))) {
stm.url<-getURL(reports$links[i])
stm.tree<-htmlTreeParse(stm.url,useInternal=TRUE )
stm.tree.parse<-unlist(xpathApply(stm.tree, path="//p", fun=xmlValue))
n<-(which(!is.na(str_locate(stm.tree.parse, "release")))+1)[1]
l<-length(stm.tree.parse)-1

# Condense separate paragraphs into one element per statement date

reports$statement.content[i]<-paste(stm.tree.parse[n:l], collapse = "")

# Remove line breaks

reports$statement.content[i]<-gsub("\r?\n|\r"," ",reports$statement.content[i])
#reports$statement.content[i]<-gsub("\\.+\\;+\\,+","",reports$statement.content[i])

# Count number of characters per statement

reports$statement.length[i]<-nchar(reports$statement.content[i])
#reports$statement.length[i]<-wordcount(reports$statement.content[i], sep = " ", count.function = sum)
}


```


```{r}
# Create R data object

saveRDS(reports, file = "fomc_data.rds")

```

```{r}

# View extracted content of a particular statement

reports$statement.dates[match(c("20070618"),reports$statement.dates)]<-"20070628"
```



```{r}
#print(reports$statement.content[2])
#nchar(reports$statement.content[2])
#print(reports$year)
#length(reports$year)
#print(reports$statement.length)
#length(reports$statement.length)
#print(reports$year, reports$statement.length, digits=1)
#print(reports[1,])

# Compute total statement length per year

yearly.length<-reports%>% group_by(year) %>% summarize(words.per.year=sum(statement.length))
yearly.length

```



```{r}
# Graph the total statement length per year

ggplot(yearly.length, aes(x=yearly.length$year,y=yearly.length$words.per.year))+geom_bar(stat="identity",fill="darkblue", colour="black") + coord_flip()+xlab("Year")+ylab("Statement Length")
```


```{r}
#Verify word count for a sample word in a sample statement

sample<-reports%>%filter(reports$statement.dates=="20140319")
sample[,4]
str_count(sample, pattern="inflation")
```


### Trend in Statement Length by year and Fed Chair
It seems that the FOMC statements became progressively verbose under Chairman Bernanke until they reached a peak in 2014 when Janet Yellen took over as the Fed Chair. Since then the statements have gotten shorter. The current chairman Jerome Powell took over in February 2018.

```{r}

# Graph the annual trend in statement length, annotated by Fed Chair

p<-ggplot(reports, aes(x=year,y=statement.length))+geom_point(stat="identity",color=statement.dates)+scale_fill_brewer(palette="Pastel1")+theme(legend.position="right")+xlab("Year") + ylab("Length of Statement")

p + ggplot2::annotate("text", x = 4,y = 5000, label = "Bernanke", family="serif", fontface="bold", colour="blue", size=4)+ggplot2::annotate("text", x=10, y=5500, label="Yellen", family="serif", fontface="bold", colour="darkred",size=4)+ggplot2::annotate("text", x=13, y=3600, label="Powell", family="serif", fontface="bold", colour="black",size=4)+ggplot2::annotate("segment", x = 0, xend = 8.1, y = 2700, yend = 6500, colour = "blue", size=1, arrow=arrow(ends="both"))+ggplot2::annotate("segment", x = 8.1, xend = 12.1, y = 6500, yend = 3200, colour = "darkred", size=1, arrow=arrow(ends="both"))+ggplot2::annotate("segment", x = 12.1, xend = 14, y = 3200, yend = 3200, colour = "black", size=1, arrow=arrow(ends="both"))
  
```


### Adding custom words and names to stop words
Remove proper nouns and irrelevant words from further analysis by adding them as custom words to the stop words lexicon

```{r}
#print(stop_words)
words<-c("committee", "ben", "geithner", "bernanke", "timothy", "hoenig", "thomas", "donald", "kevin", "mishkin", "kroszner", "kohn", "charles", "frederic")
lexicon<-c("Custom")
my.stop_words<-data.frame(words, lexicon)
colnames(my.stop_words)<-c("word","lexicon")
new.stop_words <- rbind(my.stop_words, stop_words)
new.stop_words$word<-as.character(new.stop_words$word)
new.stop_words$lexicon<-as.character(new.stop_words$lexicon)
head(new.stop_words)
```


```{r}
# Strip out punctuations, white space and custom stop words

report.words<-reports %>%mutate(date = statement.dates, year = year, text= statement.content) %>% unnest(text) %>% unnest_tokens(word, text) %>%mutate(word = stripWhitespace(gsub("[^A-Za-z ]"," ",word))) %>% filter(word != "") %>% filter(word != " ") %>%anti_join(new.stop_words)%>% count(date, year, word, sort = TRUE)%>% mutate(frequency = n) %>% select(date, year, word, frequency)
```
```{r}
# Report word frequency by statement date

report.words%>%filter(year=='2007', word=='inflation')
```


```{r fig.width = 12, fig.asp = .90}

# Rank most frequent words by year

f_text<-report.words%>% group_by(year,word) %>% summarize(total=sum(frequency))%>%arrange(year,desc(total),word)%>% mutate(rank=row_number())%>%ungroup() %>% arrange(rank,year)

#print(f_text)

topWords <- f_text %>% filter(rank<11)%>%arrange(year,rank)
print(topWords)

```


```{r fig.width = 15, fig.asp = 0.95}

# Graph top 10 most frequent words by year

gg <- ggplot(head(topWords, 130), aes(y=total,x=reorder(word,rank))) + geom_col(fill="#27408b")+
  facet_wrap(~year,scales="free", ncol=3)+ coord_flip()+theme_ridges(font_size=11)+
  labs(x="",y="",title="Most Frequent Words in FOMC Statements grouped by years (2007 - 2019")

gg
```

