---
title: "DS607_Validate_FOMC_DATA"
author: "Alexander Ng"
date: "5/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(stringr)
library(kableExtra)
library(knitr)
```

## Overview

This markdown reviews the validity of the FOMC data frame created by Jagdish.
We assume the fomc_data.rds file is located in the working directory.

```{r}
d2<-readRDS(file = "fomc_data.rds")
```

The FOMC dates are out of order and should be arrange alphabetically.
Words counts should be tabulated using *str_count*

```{r}

d2 %>% 
  mutate_if( is.factor, as.character) %>%
  arrange(statement.dates) %>%
  mutate( numwords = str_count(statement.content, "\\S+")) %>%
  select( statement.dates, numwords ) %>%
  filter(numwords > 700 ) %>% arrange(numwords)

```

## Revised DataFrame

```{r}

d2 %>% mutate_if( is.factor, as.character) %>%  # needed to change dates from factors to string
    arrange(statement.dates) %>% 
    mutate( numwords = str_count(statement.content, "\\S+")) -> d3

d3 %>% select( statement.dates, numwords) %>% 
    kable() %>% 
    kable_styling() %>% 
    scroll_box(width="85%", height="200px")

dim(d3)

```

We need to make 4 changes to the dataframe rows.

Remove 3 dates corresponding to extraordinary FOMC meeting related to swap lines, TALF and other measures.  These dates do not have normal rate setting objectives.
The dates to remove from the dataframe are:

20070810,  20080311 and 20100509.

Lastly, one of the dates is misnamed in the html filelink as 20070618 but the actual statement was released on 20070628.   This is a typo in the URL.   The actual statement contains the date June 28, 2007 which confirms the URL is misnamed.

```{r}
d3 %>% 
  filter( statement.dates != "20070810") %>%
  filter( statement.dates != "20080311") %>%
  filter( statement.dates != "20100509")  -> d4


d4[d4$statement.dates == "20070618", "statement.dates"] <- "20070628"

d4 %>% select(statement.dates, numwords) %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width="100%", height="250px")

dim(d4)

```

## Manual Validation of statements

I checked the 4 statements against the actual webpage.  

Oct 10, 2008
Dec 16, 2009
Sep 17, 2014 is the longest and is accurate.
May  1, 2019 is the most recent

They are all accurate **except for one detail**.
The elimination of newline characters is causing consecutive words to be conjoined producing nonsense words.
The line code below should be eliminated in DS607_FOMC_Sentiment_Analysis_v3.Rmd

#reports$statement.content[i]<-gsub("\n","",reports$statement.content[i])


## Conclusion

Once the removal of the minor changes to gsub command is made, the dataframe d4 above should be fit for purpose to use for research.

Nonetheless, I export the dataframe as a binary object below.

```{r}
saveRDS(d4, file = "fomc_corrected_data_v1.rds")

```
