---
title: "DS607_Merge_Classification_Data_v1"
author: "Alexander Ng"
date: "5/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(stringr)
library(lubridate)
library(kableExtra)
library(knitr)
```

## Overview

This markdown merges the corrected (of 102 x 3 data frame of FOMC statements) with the classification data.
We assume the fomc_corrected_data_v1.rds and the raw classification file is located in the working directory.



```{r}
d4<-readRDS(file = "fomc_corrected_data_v1.rds")
dim(d4)
str(d4)
```

## Raw Classification File

We explicitly override the Date column to be imported as string because we will join these two dataframes on this data.

```{r}
classificationFile = "Classification_FOMC_Statements.csv"

cls = read_csv(classificationFile , col_types = cols( Date = col_character() ) )

cls %>% rename( Economic.Growth = "Economic Growth", Employment.Growth = "Employment Growth", Medium.Term.Rate = "Medium Term Rate", Policy.Rate = "Policy Rate") -> cls

str(cls)

```

## Merging FOMC data and Classification Data

```{r}

d4 %>% inner_join( cls , by = c("statement.dates" = "Date")) %>%
  mutate( date_mdy = mdy(Date2)) %>%
  select(Index, 
         statement.dates, 
         links, 
         statement.content, 
         numwords,
         date_mdy,
         Economic.Growth,
         Employment.Growth,
         Inflation,
         Medium.Term.Rate,
         Policy.Rate ) -> mgData


str(mgData)
```

Let us view the sample data from the statements

```{r}
mgData %>% select( Index, date_mdy, Economic.Growth, Employment.Growth, Inflation, Medium.Term.Rate, Policy.Rate) %>% kable() %>% kable_styling(bootstrap_options = c("hover", "striped")) %>%
scroll_box(width = "90%", height = "300px")

```

## Exporting the Merged Data Frame

We export the merged dataframe as a single RDS object for research use.

```{r}
rds_filename = "fomc_merged_data_v1.rds"

saveRDS(mgData, file = rds_filename)
```

Let us also validate its contents.

```{r}

mgData_validated = readRDS(rds_filename)

str(mgData_validated)
```