---
title: "Text classification on the FOMC Statements"
author: "Henry Otuadinma"
date: "May 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(dplyr)
library(tidytext)
library(tm)
library(e1071)
library(R.utils)
library(DT)
library(e1071)
library(stringr)
library(lattice)
library(kernlab)
library(mlbench)
library(caretEnsemble)
library(nnet)
library(LiblineaR)
```


```{r}

fomc_data <-readRDS(file = "fomc_merged_data_v2.rds")

```


```{r}
head(select(fomc_data, Index,year,statement.dates,statement.length,date_mdy,Employment.Growth,Economic.Growth,Inflation,Medium.Term.Rate,Policy.Rate))

```


#### First, randomising the rows so that the statements from different eras of the economic movements can be well represented

```{r}

set.seed(1234567)

fomc_Rand <- fomc_data[sample(nrow(fomc_data)),]

```


##### convert the statements' textual contents to lower and remove `the federal open market committee` as it is present in all the statements

```{r}

customStopWords <- c("the federal open market committee", "committee")

```


```{r}
fomc_dataX <- fomc_Rand %>% mutate(statement.content = tolower(statement.content))

fomc_dataX$statement.content <- str_replace_all(fomc_dataX$statement.content, customStopWords, "")

```


##### Prepare the data to serve for all the classification

```{r}

# form a corpus
corpus <- VCorpus(VectorSource(fomc_dataX$statement.content))

# Remove Punctuation
corpus <- tm_map(corpus, content_transformer(removePunctuation))

# Remove numbers
corpus <- tm_map(corpus, removeNumbers)

# Convert to lower case
corpus <- tm_map(corpus, content_transformer(tolower))

# Remove stop words
corpus <- tm_map(corpus, content_transformer(removeWords), stopwords("english"))

##Stemming
corpus <- tm_map(corpus, stemDocument)

# Remove Whitespace
corpus <- tm_map(corpus, stripWhitespace)

# Create Document Term Matrix
dtm <- DocumentTermMatrix(corpus)

# handle sparsity
corpusX <- removeSparseTerms(dtm, 0.95)

# convert to matrix
data_matrix <- as.matrix(corpusX)

```


#### The `Medium.Term.Rate` variable

```{r}

mRate <- data_matrix

# attach the 'medium.term.rate' column
mRate_matrix <- cbind(mRate, fomc_dataX$Medium.Term.Rate)

# rename it to 'tone'
colnames(mRate_matrix)[ncol(mRate_matrix)] <- "tone"

# convert to data frame
mRateData <- as.data.frame(mRate_matrix)

# convert 'tone' to lower case and make it a factor column as well
mRateData$tone <- as.factor(tolower(mRateData$tone))

```


```{r}

head(mRateData[,ncol(mRateData)])
```


```{r}

mRate_n <- nrow(mRateData)
mRateTrainVolume <- round(mRate_n * 0.80)

set.seed(314)

mRateTrainIndex <- sample(mRate_n, mRateTrainVolume)
mRateTrain <- mRateData[mRateTrainIndex,]
mRateTest <- mRateData[-mRateTrainIndex,]

```


```{r}

mRateModel <- train(tone ~., data = mRateTrain, method = 'svmLinear3')

```


```{r}

mRateResult <- predict(mRateModel, newdata = mRateTest)

```


```{r}
( mRateStats = confusionMatrix( mRateResult, mRateTest$tone))
```


#### The `Economic.Growth` variable

```{r}

econGrowth <- data_matrix

# attach the 'Economic.Growth' column
econG_matrix <- cbind(econGrowth, tolower(fomc_dataX$Economic.Growth))

# rename it to 'growth'
colnames(econG_matrix)[ncol(econG_matrix)] <- "egrowth"

# convert to data frame
econData <- as.data.frame(econG_matrix)

# convert 'growth' to a factor column as well
econData$egrowth <- as.factor(econData$egrowth)

```



```{r}

econ_n <- nrow(econData)
econTrainVolume <- round(econ_n * 0.80)

set.seed(314)

econTrainIndex <- sample(econ_n, econTrainVolume)
econTrain <- econData[econTrainIndex,]
econTest <- econData[-econTrainIndex,]

```


```{r}

econModel <- train(egrowth ~., data = econTrain, method = 'svmLinear3')

```


```{r}

econResult <- predict(econModel, newdata = econTest)

```


```{r}
( econStats = confusionMatrix( econResult, econTest$egrowth))
```


#### The `Inflation` variable


```{r}

# Create Document Term Matrix
dtmI <- DocumentTermMatrix(corpus)

# handle sparsity
corpusI <- removeSparseTerms(dtm, 0.94)

# convert to matrix
data_matrixI <- as.matrix(corpusI)


inflation <- data_matrixI

# attach the 'Inflation' column
inflation_matrix <- cbind(inflation, tolower(fomc_dataX$Inflation))

# rename it to 'inflation'
colnames(inflation_matrix)[ncol(inflation_matrix)] <- "inflation"

# convert to data frame
inflationData <- as.data.frame(inflation_matrix)

# convert 'inflation' to a factor column
inflationData$inflation <- as.factor(inflationData$inflation)

```

##### remove columns that will not contribute meaninfully to the model fitting

```{r}

infDataX <- inflationData[, -which(names(inflationData) %in% c("although", "william", "william", "richard", "raphael", "randal", "san", "sarah","sandra", "togeth", "timothi" , "also","balanceconsist","basispoint","committe","dudley","elizabeth","esther","fomc"))]
 
```


```{r}

inf_n <- nrow(infDataX)
infTrainVolume <- round(inf_n * 0.70)

set.seed(314)

infTrainIndex <- sample(inf_n, infTrainVolume)
infTrain <- infDataX[infTrainIndex,]
infTest <- infDataX[-infTrainIndex,]

```


```{r}

inflationModel <- train(inflation ~., data = infTrain, method="svmLinear3")

```


```{r}

inflationResult <- predict(inflationModel, newdata = infTest)

```


```{r}
( econStats = confusionMatrix( inflationResult, infTest$inflation))
```


#### The `Employment.Growth` variable

```{r}

empGrowth <- data_matrix

# attach the 'Employment.Growth column
emp_matrix <- cbind(empGrowth, tolower(fomc_dataX$Employment.Growth))

# rename it to 'empGrowth'
colnames(emp_matrix)[ncol(emp_matrix)] <- "empGrowth"

# convert to data frame
empData <- as.data.frame(emp_matrix)

# convert 'empGrowth' to a factor column as well
empData$empGrowth <- as.factor(empData$empGrowth)

```



```{r}

emp_n <- nrow(empData)
empTrainVolume <- round(emp_n * 0.80)

set.seed(314)

empTrainIndex <- sample(emp_n, empTrainVolume)
empTrain <- empData[empTrainIndex,]
empTest <- empData[-empTrainIndex,]

```


```{r}

empModel <- train(empGrowth ~., data = empTrain, method = 'svmLinear3')

```


```{r}

empResult <- predict(empModel, newdata = empTest)

```


```{r}
( empStats = confusionMatrix( empResult, empTest$empGrowth))
```



#### The `Policy.Rate` variable

```{r}

plRate <- data_matrix

# attach the 'Policy.Rate' column
pl_matrix <- cbind(plRate, tolower(fomc_dataX$Policy.Rate))

# rename it to 'empGrowth'
colnames(pl_matrix)[ncol(pl_matrix)] <- "policy"

# convert to data frame
plData <- as.data.frame(pl_matrix)

# convert 'policy' to a factor column as well
plData$policy <- as.factor(plData$policy)

```



```{r}

pl_n <- nrow(plData)

plTrainVolume <- round(pl_n * 0.80)

set.seed(314)

plTrainIndex <- sample(pl_n, plTrainVolume)
plTrain <- plData[empTrainIndex,]
plTest <- plData[-empTrainIndex,]

```


```{r}

plModel <- train(policy ~., data = plTrain, method = 'svmLinear3')

```


```{r}

plResult <- predict(plModel, newdata = plTest)

```


```{r}
( plStats = confusionMatrix( plResult, plTest$policy))
```










